Unifying Lower Bounds on Prediction Dimension of Consistent Convex Surrogates

### 2021-05-27 cvx-flats-neurips21/cvx-flats.tex
section 4

        \begin{scope}
          \clip (h) -- (m) -- (l) -- (h);
          \foreach \c in {0,0.1,...,1}
          {
            \pgfmathsetmacro\aa{0.5*(sqrt(1-\c)+1)}
            \pgfmathsetmacro\bb{0.5*\c}
            \coordinate (A) at ($(\aa,1-\aa,0)$);
            \coordinate (B) at ($(1-\aa,\aa,0)$);
            \coordinate (C) at ($(\bb,\bb,1-2*\bb)$);
            \draw[myblue,thin] (C) parabola (A);
            \draw[myblue,thin] (C) parabola (B);
          }
          \def\c{0.6}
          \pgfmathsetmacro\aa{0.5*(sqrt(1-\c)+1)}
          \pgfmathsetmacro\bb{0.5*\c}
          \coordinate (A) at ($(\aa,1-\aa,0)$);
          \coordinate (B) at ($(1-\aa,\aa,0)$);
          \coordinate (C) at ($(\bb,\bb,1-2*\bb)$);
          \coordinate (Cleft) at ($(\bb + 2,\bb,1-2*\bb)$);
          \coordinate (Cright) at ($(\bb,\bb + 2,1-2*\bb)$);
          \coordinate (centerdot) at (0.21,0.445, 0.345);
          \draw[myblue, very thick] (C) parabola (A);
          \draw[myblue, very thick] (C) parabola (B);
          \draw[orange, dashed,shorten >= -3cm, shorten <=-3cm] (1/10,9/10,0) -- (centerdot);
          \node at (centerdot) {$\bullet$};
        \end{scope}
        \def\c{0.6}
        \pgfmathsetmacro\aa{0.5*(sqrt(1-\c)+1)}
        \pgfmathsetmacro\bb{0.5*\c}
        \node[inner sep=2pt] (levset1) at (0.45,0.55, -0.3) {\vphantom{$\Gamma_{0.1}$}};
        \node (levset) at (0.5,0.5,-.15) {$\Gamma_{0.6}$};
        \draw[-latex,thin] (levset) edge[in=330,out=180] ($(\aa,1-\aa,0)$);
        \draw[-latex,thin] (levset) edge[in=210,out=0] ($(1-\aa,\aa,0)$);
        \placefiglabel{(Q4)}
      \end{tikzpicture}
      \hspace*{-10pt}
      % \\
      % \hline
  \end{tabular}
\caption{Example properties for each quadrant. 
Throughout, we take $\bullet$ to be the distribution $p = (0.3, 0.3, 0.4)$ according to the left, top, and right outcomes respectively. 
(Q1,Q2) We cannot fit a $1$-flat (line) through $p$ without leaving the level sets $\gamma_\bot$ and $\gamma_{\text{median}}$, respectively;
Theorem~\ref{thm:convex-flats-inf-dim} implies that there is no $1$-dimensional consistent convex surrogate for either problem.
(Q3) Squared error is a $1$-dimensional convex loss, and indeed it elicits the mean of $Y$, whose level sets are all $1$-flats.
(Q4) The level sets of the variance are curved and cannot fit a $1$-flat; from Theorem~\ref{thm:convex-flats-inf-dim} there is no $1$-dimensional convex surrogate consistent for the variance.
\raf{FYI: removed some level set labels, changed $\Gamma_{0.05}$ to $\Gamma_{0.1}$ in Q3}
}\label{fig:quadrants-simplices}
\end{SCfigure}

@@

\subsection{Mode and modal interval}
\label{sec:mode-modal-interval}

\raf{Define the mode.}
\citet{dearborn2020indirect} recently showed that the mode....
\jessie{Mention it's a practical Q3 problem.}

\begin{corollary}\label{cor:mode}
  The mode has $\conscvx(\mode) = \eliccvx(\mode) = \infty$ (countably infinite) with respect to $\P$, the class of probability measures on $\Y=\reals$ which admit a smooth and bounded density and such that $\mode$ is single-valued.
\end{corollary}
\begin{proof}
  The proof of \citet[Theorem 1]{dearborn2020indirect} gives a distribution $p\in\P$ with $\mode(p) = 0 =: u$.
  It then introduces an arbitrary identification function $V:\hat\R\times\Y\to\reals^k$, $k\in\mathbb{N}$, and value $r\in\hat\R$ such that $p \in \zeros{V(r,\cdot)}$.
  Letting $F = \zeros{V(r,\cdot)}$, we therefore have an arbitrary $k$-flat containing $p$.
  The proof then proceeds to construct some $p'\in F$ with $\mode(p') \neq u$.
  Corollary~\ref{cor:Pcodim-flat-elic-relint-prop} now gives $\conscvx(\mode) \geq \eliccvx(\mode) \geq k+1$.
  \raft{PROBLEM: the set $\P$ is not convex.  Take two univariance Gaussians centered at 0 and 1: they are each in $\P$ but their mixture is not.  Do we really need $\P$ to be convex for Theorem~\ref{thm:convex-flats-inf-dim}?}
  As $k$ was arbitrary, the result follows.
\end{proof}

\raf{Introduce modal interval}
... $\Gamma_\epsilon$ is elicited by the target loss $\ell_\epsilon(r,y) = \ones\{|r-y| > \epsilon\}$
... so $\Gamma_\epsilon$ could be thought of as being in Quadrant 3.

As observed in \citet[Corollary 1]{dearborn2020indirect}, the properties $\mode$ and $\Gamma_\epsilon$ coincide with the family of distributions needed in their Theorem 1, meaning the conclusion of Corollary~\ref{cor:mode} transfers to the modal interval as well.
\begin{corollary}\label{cor:modal-interval}
  For any $\epsilon > 0$, the modal midpoint, $\Gamma_\epsilon: \P_\epsilon \to \reals$, has  $\conscvx(\Gamma_\epsilon) = \eliccvx(\Gamma_\epsilon) = \infty$ (countably infinite) with respect to $\P_\epsilon$, the class of unimodal probability measures on $\Y=\reals$ which admit a smooth and bounded density, and such that $\mode$ and $\Gamma_\epsilon$ are single-valued.
\end{corollary}

Thus, while $\Gamma_\epsilon$ is elicitable, it does not have any consistent finite-dimensional convex surrogate.
\raf{Discussion of why this sort of makes sense as a limit of 0-1 loss and the mode.}

>>>>

        \begin{scope}
          \clip (h) -- (m) -- (l) -- (h);
          \foreach \c in {0,0.1,...,1}
          {
            \pgfmathsetmacro\aa{0.5*(sqrt(1-\c)+1)}
            \pgfmathsetmacro\bb{0.5*\c}
            \coordinate (A) at ($(\aa,1-\aa,0)$);
            \coordinate (B) at ($(1-\aa,\aa,0)$);
            \coordinate (C) at ($(\bb,\bb,1-2*\bb)$);
            \draw[myblue,thin] (C) parabola (A);
            \draw[myblue,thin] (C) parabola (B);
          }
          \def\c{0.6}
          \pgfmathsetmacro\aa{0.5*(sqrt(1-\c)+1)}
          \pgfmathsetmacro\bb{0.5*\c}
          \coordinate (A) at ($(\aa,1-\aa,0)$);
          \coordinate (B) at ($(1-\aa,\aa,0)$);
          \coordinate (C) at ($(\bb,\bb,1-2*\bb)$);
          \coordinate (Cleft) at ($(\bb + 2,\bb,1-2*\bb)$);
          \coordinate (Cright) at ($(\bb,\bb + 2,1-2*\bb)$);
          \coordinate (centerdot) at (0.21,0.445, 0.345);
          \draw[myblue, very thick] (C) parabola (A);
          \draw[myblue, very thick] (C) parabola (B);
          \draw[orange, dashed,shorten >= -3cm, shorten <=-3cm] (1/10,9/10,0) -- (centerdot);
          \node at (centerdot) {$\bullet$};
        \end{scope}
        \def\c{0.6}
        \pgfmathsetmacro\aa{0.5*(sqrt(1-\c)+1)}
        \pgfmathsetmacro\bb{0.5*\c}
        \node[inner sep=2pt] (levset1) at (0.45,0.55, -0.3) {\vphantom{$\Gamma_{0.1}$}};
        \node (levset) at (0.5,0.5,-.15) {$\Gamma_{0.6}$};
        \draw[-latex,thin] (levset) edge[in=330,out=180] ($(\aa,1-\aa,0)$);
        \draw[-latex,thin] (levset) edge[in=210,out=0] ($(1-\aa,\aa,0)$);
        \placefiglabel{(Q4)}
      \end{tikzpicture}
      \hspace*{-10pt}
      % \\
      % \hline
  \end{tabular}
\caption{Example properties for each quadrant. 
Throughout, we take $\bullet$ to be the distribution $p = (0.3, 0.3, 0.4)$ according to the left, top, and right outcomes respectively. 
(Q1,Q2) We cannot fit a $1$-flat (line) through $p$ without leaving the level sets $\gamma_\bot$ and $\gamma_{\text{median}}$, respectively;
Theorem~\ref{thm:convex-flats-inf-dim} implies that there is no $1$-dimensional consistent convex surrogate for either problem.
(Q3) Squared error is a $1$-dimensional convex loss, and indeed it elicits the mean of $Y$, whose level sets are all $1$-flats.
(Q4) The level sets of the variance are curved and cannot fit a $1$-flat; from Theorem~\ref{thm:convex-flats-inf-dim} there is no $1$-dimensional convex surrogate consistent for the variance.
-}\label{fig:quadrants-simplices}
\end{SCfigure}

@@

\subsection{Mode and modal interval (Q4, Q3)}
\label{sec:mode-modal-interval}

For finite $|\Y|$, the mode $\mode(p) = \argmax_{y\in\Y} p(y)$ is elicited by 0-1 loss.
By contrast, for $\Y=\reals$, the mode is not elicitable~\cite{heinrich2013mode}, landing it in Quadrant 4.
Defining the mode is subtle for general distributions; here let us assume $p$ has a smooth and bounded Lesbegue density $f_p$, and define the mode the same way, $\mode(p) = \argmax_{y\in\Y} f_p(y)$.
\citet{dearborn2020indirect} recently showed a strong impossibility result, that the mode has countably infinite elicitation complexity with respect to identifiable properties.
In other words, it is as hard to elicit the mode as the full distribution $p$ itself.
Complexity with respect to convex-elicitable properties is left as an important open question.

We resolve this question, with a matching infinite lower bound for convex-elicitable properties.
In light of our $d$-flats framework, the result is nearly immediate, as the proof in \citet{dearborn2020indirect} already showed that the level sets of the mode cannot contain any $d$-flats.

\begin{theorem}\label{thm:mode}
  The mode has $\conscvx(\mode) = \eliccvx(\mode) = \infty$ (countably infinite) with respect to $\P$, the class of probability measures on $\Y=\reals$ which admit a smooth and bounded density and such that $\mode$ is single-valued.
\end{theorem}
\begin{proof}
  The proof of \citet[Theorem 1]{dearborn2020indirect} gives a distribution $p\in\P$ with $\mode(p) = 0 =: u$.
  It then introduces an arbitrary identification function $V:\hat\R\times\Y\to\reals^k$, $k\in\mathbb{N}$, and value $r\in\hat\R$ such that $p \in \zeros{V(r,\cdot)}$.
  Letting $F = \zeros{V(r,\cdot)}$, we therefore have an arbitrary $k$-flat containing $p$.
  The proof then proceeds to construct some $p'\in F$ with $\mode(p') \neq u$.
  Corollary~\ref{cor:Pcodim-flat-elic-relint-prop} now gives $\conscvx(\mode) \geq \eliccvx(\mode) \geq k+1$.
  \raft{PROBLEM: the set $\P$ is not convex.  Take two univariance Gaussians centered at 0 and 1: they are each in $\P$ but their mixture is not.  Do we really need $\P$ to be convex for Theorem~\ref{thm:convex-flats-inf-dim}?}
  As $k$ was arbitrary, the result follows.
\end{proof}

A closely related property for any $\beta > 0$ is the (midpoint of the) modal interval of width $2\beta$, given by $\Gamma_\beta(p) = \argmax_{x\in\reals} p([x-\beta,x+\beta])$.
Interestingly, unlike the mode for $\Y=\reals$, the modal interval is elicitable, by the target loss $\ell_\beta(r,y) = \ones\{|r-y| > \beta\}$.
The problem of estimating the modal interval therefore could be thought of as falling under Quadrant 3.

As observed in \citet[Corollary 1]{dearborn2020indirect}, the properties $\mode$ and $\Gamma_\beta$ coincide with the family of distributions needed in their Theorem 1, meaning the conclusion of Theorem~\ref{thm:mode} transfers to the modal interval as well.
\begin{corollary}\label{cor:modal-interval}
  For any $\beta > 0$, the modal midpoint, $\Gamma_\beta: \P_\beta \to \reals$, has  $\conscvx(\Gamma_\beta) = \eliccvx(\Gamma_\beta) = \infty$ (countably infinite) with respect to $\P_\beta$, the class of unimodal probability measures on $\Y=\reals$ which admit a smooth and bounded density, and such that $\mode$ and $\Gamma_\beta$ are single-valued.
\end{corollary}

Thus, while $\Gamma_\beta$ is elicitable, it does not have any consistent finite-dimensional convex surrogate.
\raf{Discussion of why this sort of makes sense as a limit of 0-1 loss and the mode.}
